---
title: 'Panel in R: interact time FE with space units'
author: "Matthieu"
abstract: "Goal is to add interactions with a time effects and some group, such as region, county etc"
date: "May 15, 2019"
output:
  html_document:
    toc: true
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(plm)
library(lfe)
library(broom)
library(tidyverse)
```


# Summary, refs

Discussed on Stack Overflow:

- [r-plm-and-lm-fixed-effects](https://stackoverflow.com/questions/43636724/r-plm-and-lm-fixed-effects)
- [fixed-effects-regression-with-state-specific-trends](https://stackoverflow.com/questions/34232834/fixed-effects-regression-with-state-specific-trends)

There is a quite complicated story about counting the degrees of freedom: 

- standard FE2: just need to remove one df
- year-region: here interacting time with 9 regions implies we should remove (i.e. normalize the parameters) $n_R$ -1=8 here!

`lm()` and `felm()` give same answer (-8), `plm()` not. Getting the fixed-effects in `felm()` shows indeed 8 FE are set to 0. In `plm()`, noen of them is zero, so it is not clear what the transformation is?



State specific trend: need to convert to integer

```{r, eval=FALSE}
fe1_Ttrend_plm <- plm(log(gsp) ~  state : as.integer(year), data = produc_plm)
```


# Preliminary: Load data

```{r}
data("Produc", package = "plm")
```

Add region:

```{r}
Produc$year_region <- paste(Produc$year, Produc$region, sep = "_")
```


# With package plm

## Simple reg

```{r}
zz <- plm(gsp ~ pcap, data = Produc, index = c("state","year"), effect = "twoway")
summary(zz)
```

## Reg with year_region

```{r}
zz2 <- plm(gsp ~ pcap, data = Produc, index = c("state","year_region"), effect = "twoway")
zz
zz2
```


Check FE2:

```{r}
FE_1 <- fixef(zz, effect = "time")
FE_2 <- fixef(zz2, effect = "time")
FE_df_1 <- tibble(fe = names(FE_1), value = FE_1) %>% 
  mutate(year = as.integer(fe))
FE_df_2 <- tibble(fe = names(FE_2), value = FE_2) %>% 
  separate(fe, c("year", "region"), remove = FALSE, convert = TRUE)
FE_df_2
```


Average FE

```{r}
## ave
FE_df_2_ave <- FE_df_2 %>% 
  group_by(year) %>% 
  summarise(value = mean(value)) %>% 
  ungroup()

FE_df_both <- rbind(select(FE_df_1, -fe),
                    FE_df_2_ave) %>% 
  mutate(type = rep(c("FE_1", "FE_2"), each = nrow(FE_df_1)))

FE_df_both
```


Plot it
```{r}
## plot
FE_df_both %>% 
  ggplot(aes(x = year, y = value, colour = type))+
  geom_point()
```




# With lm

```{r}
lm1 <- lm(gsp ~ pcap -1 + state + factor(year), data = Produc)
lm2 <- lm(gsp ~ pcap -1 + state + year_region, data = Produc)
```

Compare:

```{r}
all.equal(tidy(lm1) %>% 
            filter(term=="pcap") %>% 
            as.data.frame(), 
          tidy(zz) %>%
            as.data.frame())

all.equal(tidy(lm2) %>% 
            filter(term=="pcap") %>% 
            as.data.frame(), 
          tidy(zz2) %>%
            as.data.frame())
```



# With package lfe (felm  function)

### Run simple

```{r}
## year
felm <- felm(gsp ~ pcap|state+year, data = Produc)
all.equal(tidy(felm) %>% as.data.frame(), 
          tidy(zz)%>% as.data.frame())

```

### Run with year_region

```{r}
## year region
felm2 <- felm(gsp ~ pcap|state+year_region, data = Produc, exactDOF="rM")

all.equal(coef(felm2), coef(zz2))

## but SE not the same!
all.equal(tidy(felm2) %>% as.data.frame(), 
          tidy(zz2)%>% as.data.frame())

## but same as lm!?
all.equal(tidy(lm2) %>% 
            filter(term=="pcap") %>% 
            as.data.frame(), 
          tidy(felm2) %>%
            as.data.frame())
```

Why not same? Df are different! felm has 8 more!

```{r}
c(felm$df.residual, zz$df.residual)
c(felm=felm2$df.residual, plm=zz2$df.residual)
diff(c(felm2$df.residual, zz2$df.residual))
```

```{r}
Produc %>% 
  summarise_at(c("state", "year", "year_region", "region"), n_distinct)

## so technically, we estimate 48+153 = 201
nrow(Produc) - 201

## This is the df of plm!

## But many FEs are zero!
getfe(felm) %>% 
  count(effect_zero=effect==0)

getfe(felm2) %>% 
  count(effect_zero=effect==0)


## plm 
tibble(n_ind =length(fixef(zz, effect = "individual")),
       n_time =length(fixef(zz, effect = "time")), 
       tot = n_ind+n_time)

tibble(n_ind =length(fixef(zz2, effect = "individual")),
       n_time =length(fixef(zz2, effect = "time")), 
       tot = n_ind+n_time)
```



Do with factors:

```{r}
Produc2 <- Produc %>% 
  mutate(region = as.factor(region),
         year = as.factor(year))
felm2b <- felm(gsp ~ pcap|state+year:region, data = Produc2)

all.equal(tidy(felm2) %>% as.data.frame(), 
          tidy(felm2b)%>% as.data.frame())
```



