---
title: "vars Impulse response fucntion with ggplot2"
author: "Matthieu"
date: "January 10, 2019"
output:
  html_document:
    keep_md: yes
---


## Objective

using package `vars`, the function `irf()` produces Impulse Response Function. I wanted to plot them with `ggplot2`, instead of the built-in `plot()` function. This script provides two functions:

* `as.data.frame.varirf()` converts output intoa  standard long data.frame
* `irfplot()` : do the whole conversion and plotting

## Example


```{r, message= FALSE}
library(tidyverse)
library(devtools)
source_url("https://raw.githubusercontent.com/MatthieuStigler/Misconometrics/master/irf_for_ggplot/as.data.frame.varirf.R")

## standard VAR analysis
library(vars)
data(Canada)
var.2c <- VAR(Canada, p = 2, type = "const")
irf_1 <- irf(var.2c, runs = 1000)
irf_1_rw <- irf(var.2c, impulse = "rw", response =  "rw", boot = TRUE, runs = 500)
```


Plot now all

```{r, message= FALSE}
irfplot(irf_1)

```

Plot when called with just 1:

```{r}
irfplot(irf_1_rw)
```

Check various options of the as.data.frame method:

```{r}
as.data.frame(x=irf_1, format = "long") %>%  as_tibble()
as.data.frame(x=irf_1, format = "wide_resp") %>%  as_tibble()
as.data.frame(x=irf_1, format = "wide_type") %>%  as_tibble()
```

  


