---
title: "Prospect functions: illustration"
author: "Mat"
date: "11/5/2020"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Goal

I am just trying to understand the intuition behind the cumulative prospect theory: what does it mean to: 

1. Use rank-dependent utility
2. Use different weights for gain/losses and a different formulation of the probabilities? 

Case study: fair dice with equi-probability 1/6. Values 1 to 3 lead to a loss, values 4:6 to a gain. This is as in Tversky-Kahneman (1993), although for now I don't consider differences the outcomes they attribute to 1 to 6.

```{r, message=FALSE}
library(tidyverse)
```


## Rank dependent utility


```{r, fig.width=12, fig.height=5}
weight_S <- function(p, gam=0.61) p^gam /(p^gam +(1-p)^gam)^(1/gam)
df_rank_depend <- tibble(outcome=1:6,
       prob = 1/6,
       prob_cum = rev(cumsum(prob)),
       prob_cum_lead = dplyr::lead(prob_cum, default=0),
       dec_convex= prob_cum^1.2-prob_cum_lead^1.2,
       dec_concave= prob_cum^0.5-prob_cum_lead^0.5,
       dec_linear= prob_cum-prob_cum_lead,
       dec_S_shape=weight_S(prob_cum)-weight_S(prob_cum_lead))

df_rank_depend%>% 
  pivot_longer(starts_with("dec_"), names_prefix="dec_", names_to="weighting")  %>% 
  mutate(type=weighting=="linear") %>% 
  ggplot(aes(x=outcome, y=value, color=weighting))+
  geom_line(aes(linetype=type))+
  scale_x_continuous(breaks=1:6)+
  xlab("Dice outcome")+ylab("Weight")+
  scale_linetype(guide = 'none')+
  ggtitle("Rank-dependent utility with various weighting functions")

```

## Prospect theory

```{r, fig.width=12, fig.height=5}
weight_S_minus <- function(p, gam=0.69) p^gam /(p^gam +(1-p)^gam)^(1/gam)

df_prospect <- tibble(outcome=1:6,
       gain=c(-5, -3, -1, 2, 4, 6),
       prob = 1/6,
       prob_cum = if_else(gain<0, cumsum(prob), rev(cumsum(prob))),
       prob_cum_lead = if_else(gain<0,
                              dplyr::lag(prob_cum, default=0),
                              dplyr::lead(prob_cum, default=0)),
       dec_linear = prob_cum-prob_cum_lead,
       dec_S_sym = weight_S(prob_cum)-weight_S(prob_cum_lead),
       dec_S_asym =  if_else(gain<0,
                             weight_S_minus(prob_cum)-weight_S_minus(prob_cum_lead),
                             weight_S(prob_cum)-weight_S(prob_cum_lead))) %>% 
  mutate(dec_S_shape_rank_depend=df_rank_depend$dec_S_shape) 

df_prospect%>% 
  pivot_longer(starts_with("dec_"), names_prefix="dec_", names_to="weighting")  %>% 
  mutate(type=weighting=="linear") %>% 
  ggplot(aes(x=outcome, y=value, color=weighting))+
  geom_line(aes(linetype=type))+
  scale_x_continuous(breaks=1:6)+
  xlab("Dice outcome")+ylab("Weight")+
   scale_linetype(guide = 'none')+
  geom_vline(xintercept = 3.5, linetype=2)+
  ggtitle("Cumulative prospect utility with various weighting functions")
```

### Cumulative used

```{r}
df_probs <- df_rank_depend %>% 
  select(outcome, starts_with("prob_cum")) %>% 
  mutate(type="rank dependent theory") %>% 
  rbind(df_prospect %>% 
          select(outcome, starts_with("prob_cum")) %>% 
          mutate(type="cumul prospect theory")) %>% 
  mutate(type=fct_relevel(type, "rank dependent theory"))

df_probs%>% 
  ggplot(aes(x=outcome, y=prob_cum))+
  geom_line()+
  geom_line(aes(y=prob_cum_lead), linetype=2)+
  facet_wrap(~type)+
  ggtitle("Comparing cumulative probabilities")+
  geom_vline(aes(xintercept=pos),
             data =tibble(pos=3.5,
                          type = factor("cumul prospect theory", levels = levels(df_probs$type))),
             linetype=2, col="blue")
```




```{r, include=FALSE}
   # add this chunk to end of mycode.rmd
   file.rename(from="prospect_utility_VISU.md", 
               to="README.md")
```


