---
title: 'dec_covar: R implementation of Gelbachs covariate decomposition: `When Do Covariates Matter? And Which Ones, and How Much?`'
author: "Matthieu"
date: "September 4, 2018"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

This code implements the basics of the paper: Gelbach (2016) *When Do Covariates Matter? And Which Ones, and How Much?* Journal of Labor Economics, 2016, vol. 34(2)

It does just do the decomposition, without inference. It does not do the grouping either: in the standard OLS case, R can do a multivariate OLS, so computing the auxiliary regressions for one or multiple LHS variables is done efficiently. User can simply use `tidyverse::group_by` tools to do the grouping. 

Underlying `dec_covar` is the function `reg_aux`, which regresses all covariates in a OLS/panel onto the main regressor of interest. This function might be of use per se, as it is written in an efficient manner: it uses lm and felm capacity to handle multi-response models, and also implements the sweep operator, which can be much faster with large datasets!


# Use

The function is: `dec_covar(object = , var_main)`. `var_main` is the variable of interest, as character string.

The function works for `object` of class:

 - OLS regressions computed with `lm()` 
 - panel computed with  `plm()` or `felm()` from package *lfe*

# Citation

If you use this script in an academic publication, I would appreciate if you could cite it accordingly:

Stigler, Matthieu (2018) "dec_covar, an R implementation of Gelbach's covariate decomposition". https://github.com/MatthieuStigler/Misconometrics/tree/master/Gelbach_decompo


# Illustration


1) Source the script:

```{r cars, message = FALSE}
devtools::source_url("https://raw.githubusercontent.com/MatthieuStigler/Misconometrics/master/Gelbach_decompo/dec_covar.R")
```

2) Run your OLS model, with all the variables. We use here built-in data *swiss*:

```{r}
model_full_1 <- lm(Fertility ~ ., data=swiss)
```

3) Use the function `dec_covar` on the OLS output, specify which variable is the *main* one:

```{r}
dec <- dec_covar(object = model_full_1, var_main = "Education")
```



```{r, echo= FALSE}
library(knitr)
kable(dec, format = "markdown")
```

This will give you the impact of each coefficient on the comparison between the regression with all covariates and the one with only `Education`

```{r}
model_base <- lm(Fertility ~ Education, data=swiss)
coef(model_base)["Education"]-coef(model_full_1)["Education"]
subset(dec, covariate=="Total")$delta_Education
```


## Plots

You can get two plots also. You will need to use the `format="long"` argument, as well as `conf.int=TRUE` (that's just for the beta and gamma though)

```{r}
dec_long <- dec_covar(object = model_full_1, var_main = "Education", format = "long", add_coefs = TRUE, conf.int = TRUE)
```

```{r plot_dec}
plot_dec(dec_long) +
  ggtitle("Effect of each covariate on the main variable's coef")
```

```{r plot_gamma_beta}
plot_gamma_beta(dec_long, add_CI = TRUE) +
  ggtitle("Covariate impact: direct (beta) and indirect (gamma) impact")
```


```{r plot_gam_bet_del}
plot_gam_bet_del(dec_long)
```

