---
title: "Anova manual"
author: "Matthieu"
date: "December 17, 2019"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Goal


Goal of this quick note is to do a manual ANOVA, and see how to relate different levels of factors. 

Facts learned:

 - When a factor is subset of another, `anova` will drop larger factor (cf region and state)
 - If factor orthogonal, variance is additive
 - With two hierarchical factors, doing an anova on the residuals of the higher factor will give same results. 

## Prepare data and models
 
```{r, message=FALSE}
library(plm)
library(tidyverse)
library(matPkg)

data(Produc)
Produc <- as_tibble(Produc) %>% 
  mutate(zone = if_else(region %in% c("1", "2", "3"), "A", "B"))
```
 
 
Estimate models:

```{r}
tot_SSR <- var(Produc$unemp)*(nrow(Produc)-1)


lm0 <- lm(unemp~1, data=Produc)
lm1_y <- lm(unemp~year, data=Produc)
lm1_s <- lm(unemp~state, data=Produc)
lm1_r <- lm(unemp~region, data=Produc)
lm1_z <- lm(unemp~zone, data=Produc)
lm2_ys <- lm(unemp~year+state, data=Produc)
lm2_yr <- lm(unemp~year+region, data=Produc)
lm2_sr <- lm(unemp~state+region, data=Produc)
lm3 <- lm(unemp~year+state+region, data=Produc)
```
 
Estimate cumulative models
```{r}
lm_cum_sr <- lm(unemp~region, data=mutate(Produc, unemp=predict(lm1_s)))
```
 

## Convenience functions

```{r}
aov_clean <- function(x) {
  nam <- rlang::ensym(x)
  as.data.frame(anova(x)) %>% 
    rownames_to_column(var = "variable") %>% 
    rename(SSR = `Sum Sq`) %>% 
    mutate(SSR_perc = 100* SSR/sum(SSR)) %>% 
    mat_add_total_row() %>% 
    mutate(model = as.character(nam)) %>% 
    select(model, variable, Df, SSR, SSR_perc)
  
}

aov_alter <- function(x) {
  nam <- rlang::ensym(x)
  
  res <- residuals(x)
  tibble(model = as.character(nam),
         variable ="Residuals",
         SSR = var(res) * (nobs(x)-1),
         SSR_perc = 100*SSR/tot_SSR)
}

```

## Check results

```{r}
aov_clean(x=lm0)
aov_clean(x=lm1_y)
aov_clean(x=lm1_s)
aov_clean(x=lm1_r)
aov_clean(x=lm2_ys)
aov_clean(x=lm2_yr)
aov_clean(x=lm2_sr)
aov_clean(x=lm3)
```

### Alternative: residuals

Check with the alter method: just var of residuals

```{r}
#
aov_clean(x=lm2_yr)
aov_alter(x=lm2_yr)
```

### Decompose hierarchical ones

Check residual approach: if state gives 31%, and region 16%, can I claim that region is generating 16/31 of variance? yes

```{r}
aov_clean(lm1_s)
aov_clean(lm1_r)
aov_clean(lm_cum_sr)
```

Using residuals only:

```{r}
cumdiff <- function(x) cumsum(diff(x))

dat <- rbind(aov_alter(lm1_z) %>% 
               mutate(variable="resid_zone"),
             aov_alter(lm1_s) %>% 
               mutate(variable="resid_state"),
             aov_alter(lm1_r)%>% 
               mutate(variable="resid_region")) %>% 
  bind_rows(tibble(model = "all",
                   variable ="resid_total",
                   SSR = tot_SSR, SSR_perc=100))

dat
dat[1:3,] %>% 
  mutate(var_take = 100 - SSR_perc) %>%   
  arrange(var_take) %>% 
  mutate(var_use = c(var_take[1], abs(diff(var_take))),
         check = cumsum(var_use)) %>% 
  as.data.frame()

```

Check: want to have same values as in column `var_take`:

```{r}
rbind(aov_clean(lm1_z),
      aov_clean(lm1_r),
      aov_clean(lm1_s)) %>% 
  filter(!variable %in% c("Total", "Residuals"))
```

